---
title: "Boidr analysis"
output:
  pdf_document:
    highlight: tango
---

Load libraries, install boidr
```{r}
library(tidyverse)
library(tidyr)
library(ggplot2)
library(tictoc)
# there is a bug, need to be installed from github to use the latest version: remotes::install_github("extendr/rextendr")
library(rextendr)
setwd("~/Source/Repos/boids_rust/boidr")

rextendr::document()
devtools::load_all(".")
# this is loud
invisible({capture.output({

})})

# data_folder <- "~/Source/Repos/boids_rust/boidranalysis/Data/"
data_folder <- "./Data/"
```
```{r}
source("./R/helpers.R")
source("./R/directions_angles.R")
source("./R/flock_metrics.R")
```

get flocking data
```{r}
# boid_data <- flock_return(no_iter = 8000, init_boids = 256, save_locations_path = data_folder, sample_rate = 8, init_width = 4000, init_height = 4000)
# boid_data <- flock_detailed(no_iter = 8000,
#                              init_boids = 512,
#                              save_locations_path = "", # data_folder,
#                              sample_rate = 32,
#                              init_width = 4000, 
#                              init_height = 4000, 
#                              sensory_distance = 50,
#                              allignment_coef = .02,
#                              allignment_trs_coef = 1.15,
#                              cohesion_coef = 0.002, 
#                              cohesion_trs_coef = .95,
#                              separation_coef = 4.1,
#                              separation_trs_coef = .3,
#                              min_speed = .5,
#                              max_speed = 2.,
#                              max_steering = .65,
#                              dbscan_clustering = T,
#                              boundary_config = "{\"type\": \"Thoroidal\"}"
#                              # boundary_config = "{\"type\": \"Repulsive\", \"distance\": 100, \"force\": 0.05}"
#                             ) %>%
#     tibble()

tic()
config <- get_config(
  # "string_ball_hybrid.toml", 
  # "large_string.toml", 
  "smaller_string.toml", 
    overwrite = list(
      init_boids = 2^10,
      no_iter = 2^17,
      init_width = 1600, 
      init_height = 1600,
      min_speed = 2,
      max_speed = 2,
      sample_rate = 32,
      wander_coef = .1,
      # boundary_config = "{\"type\": \"Repulsive\", \"distance\": 100, \"force\": 0.05}"
      boundary_config = "{\"type\": \"Toroidal\"}",
      distance_config = "{\"type\": \"EucToroidal\"}"
      # field_of_vision = 360.
    )
)
boid_data <- rlang::exec(flock_detailed, !!!config) %>%
    tibble()
toc()
```

```{r}
tic("get_directional_boid_data_2 on 400MB data")
direction_data_by_boid <-  read_csv("Data/0314_experiment2_h1/boids-data_1678856512443.csv") |>
  get_directional_boid_data2()
toc()
tic("get_directional_boid_data_2 on 400MB data")
direction_data_by_boid <-  read_csv("Data/0307_experiment2_e1_1/boids-data_1678166144706.csv") |>
  get_directional_boid_data2()
toc()


tic("get_directional_boid_data_2 on 141MB data")
direction_data_by_boid <-  read_csv("Data/0301_experiment2_c2/boids-data_1677727700033.csv") |>
  get_directional_boid_data2()
toc()

tic()
preprocess_file("./Data/0301_experiment2_c2/boids-data_1677727700033.csv", init_width = 8000, init_height = 8000, no_boids = 8192)

file.remove("./Data/0301_experiment2_c2/boids-data_1677727700033.csv")
direction_data_by_boid <- read_csv("./Data/0301_experiment2_c2/prepro_boids-data_1677727700033.csv") |>
  get_directional_boid_data3()
toc()
```

visualize
```{r}
# sub_sub_sample_rate <- 124
# # direction_data_by_boid <- boid_data %>%
# direction_data_by_boid <- boid_data %>%
#   # sub_sample_data(every_nth = sub_sub_sample_rate) %>%
#   get_directional_boid_data2(remove_boundary = FALSE)
# 
# print("range of bearings:")
# range(direction_data_by_boid$bearings)
# 
# print("range of headings:")
# range(direction_data_by_boid$headings)

test_time <- 5

headings_binned <- direction_data_by_boid %>%
  filter(time <= test_time) %>%
  mutate(
    headings = floor(rad2deg(headings))
  ) %>%
  group_by(headings) %>%
  summarise(
    count = length(id)
  ) %>%
  select(heading = headings, count = count)

bearings_binned <- direction_data_by_boid %>%
  filter(time <= test_time) %>%
  mutate(
    bearings = floor(rad2deg(bearings))
  ) %>%
  group_by(bearings) %>%
  summarise(
    count = length(id)
  ) %>%
  select(bearing = bearings, count = count)

summary(direction_data_by_boid %>% select(-id))
```

```{r}
ggplot(headings_binned, aes(x = heading, y = log(count))) +
  coord_polar(theta = "x", start = pi) +
  geom_bar(stat = "identity", fill = "orange", width = .9) +
  labs(x = "Heading w.r.t N") +
  theme_bw()
ggplot(headings_binned, aes(x = heading, y = count)) +
  coord_polar(theta = "x", start = pi) +
  geom_bar(stat = "identity", fill = "orange", width = .9) +
  labs(x = "Heading w.r.t N") +
  theme_bw()
```

```{r}
ggplot(bearings_binned, aes(x = bearing, y = log(count))) +
  coord_polar(theta = "x", start = pi) +
  geom_bar(stat = "identity", fill = "deeppink4", width = .9) +
  labs(x = "Turning angle") +
  theme_bw()
ggplot(bearings_binned, aes(x = bearing, y = count)) +
  coord_polar(theta = "x", start = pi) +
  geom_bar(stat = "identity", fill = "deeppink4", width = .9) +
  labs(x = "Turning angle") +
  theme_bw()
```
The symmetry here is fascinating, after investigation I am confident in the bearings function. A naive explanation would be the boid's tendency to over react and then over correct, e.g., when it wants to avoid a boid in front of of him slightly on the left, it will turn abruptly to the right, then starts to be attracted again towards the same boid.
```{r}
ggplot(direction_data_by_boid, aes(x=headings)) +
    # coord_polar(theta = "x", start = pi) +
geom_density(color="aquamarine4", fill="aquamarine3") +
  theme_bw()
```
```{r}
ggplot(direction_data_by_boid, aes(x=bearings)) +
  geom_density(color="aquamarine4", fill="aquamarine3") +
  theme_bw()
```
// note: high sample_rate > 512, start seeing "swan tailing" at the tails

```{r}
avg_norm_data <- get_average_norm_vel(direction_data_by_boid)


  ggplot(
    data = tibble(
      avg_norm_vel = avg_norm_data$avg_norm_vel,
      t = avg_norm_data$time * config$sample_rate * 32# * sub_sub_sample_rate
    ),
    aes(x = t, y = avg_norm_vel)
  ) +
    geom_line() +
    theme_bw() +
    labs(x = "time t", y = "mean average norm vel",
         title = paste(
           "average norm vel;",
           "agents:", config$init_boids, ";",
           "iterations:", config$no_iter, ";",
           "sample:", config$sample_rate, ";"
         )
    )
```
```{r}
direction_data_by_boid %>%
  filter(cluster_id != 0) %>%
  select(dx, dy, time, cluster_id) %>%
  group_by(time) %>%
  group_by(time, cluster_id) %>%
  summarise(
    v0x = sum(abs(dx)) / n(),
    v0y = sum(abs(dy)) / n(),
    vix = abs(sum(dx)),
    viy = abs(sum(dy)),
    count = n()
  ) %>%
  mutate(result = sqrt(vix^2 + viy^2) / (count * sqrt(v0x^2 + v0y^2))) %>%
  ggplot(aes(y = result)) +
geom_point(aes(x = time, group = cluster_id, color = result))
```

```{r}

ggplot(headings_binned) +
  geom_col(aes(x = heading, y = log(count))) +
  scale_x_continuous(
    breaks = seq(-180, 179, 45),
    minor_breaks = seq(-180, 179, 15)
  ) +
  # coord_polar() +
  coord_polar(theta = "x", start = pi) +
  theme_bw()
ggplot(headings_binned) +
  geom_col(aes(x = heading, y = count)) +
  scale_x_continuous(
    breaks = seq(-180, 179, 45),
    minor_breaks = seq(-180, 179, 15)
  ) +
  # coord_polar() +
  coord_polar(theta = "x", start = pi) +
  theme_bw()
```
```{r}

sensory_distance = max(
  max(config$cohesion_trs_coef * config$sensory_distance,
      config$separation_trs_coef * config$sensory_distance)
  ,config$allignment_trs_coef * config$sensory_distance)

pi * sensory_distance^2


test <- direction_data_by_boid %>%
  group_by(time) %>%
  reframe(voronoi_areas = list(get_voronoi_areas(x, y, 4000, 4000)))

test %>%
  rowwise() %>%
  mutate(time, C_voronoi = sum(voronoi_areas < pi * sensory_distance^2) / 2048) %>%
  select(time, C_voronoi) %>%
  ggplot(aes(x = time, y = C_voronoi)) +
  geom_point()
```

```{r}
direction_data_by_boid %>%
  arrange(time) %>%
  group_by(id) %>%
  mutate(breaings_ma=rollapply(bearings, 5,mean,align='right',fill=NA)) %>%
  ungroup() %>%
  group_by(time, cluster_id) %>%
  summarise(breaings_ma = mean(breaings_ma)) %>%
  ungroup() %>%
  filter(!is.na(breaings_ma)) %>%
  mutate(noise = cluster_id == 0) %>%
  ggplot(aes(x = time, y = breaings_ma)) +
  geom_point(size = .5, alpha = .1, aes(color = noise))  +
  stat_summary(aes(y = breaings_ma, color = noise), fun=mean, geom="line") +
  theme_bw()
```
```{r}


```
